---
title: "r_solo-bdfinlayson"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(stringr)
library(ggplot2)
```

## 1. Read in `districts.csv` into a tibble named `districts`. How many rows and columns does it contain?

```{r}

districts <- read_csv('./data/districts.csv')

nrows <- districts %>% 
  count()

ncols <- districts %>% 
  ncol()

str_interp("There are ${nrows} rows and ${ncols} columns")

```


## 2. Notice that the first row corresponds to the whole State of Tennessee. Remove this row and save the result back to `districts`.

```{r}

districts <- districts[-1,]

districts %>% head()
```

## 3. How many districts have a proficiency rate of at least 80% for both alg_1 and eng_1?

```{r}

ndistricts_proficiency_80 <- districts %>%
  filter(alg_1 >= 80 & eng_1 >= 80) %>% 
  summarize(n_distinct(system))

str_interp("There are ${ndistricts_proficiency_80} districts with alg and eng proficiency greater than 80%")
```

## 4. How many districts have a proviciency rate less than 50% for either alg_1 or eng_1?

```{r}
ndistricts_proficiency_50 <- districts %>% 
  filter(alg_1 < 50 | eng_1 < 50) %>% 
  summarise(n_distinct(system))

str_interp("There are ${ndistricts_proficiency_50} districts with alg and eng proficiency less than 50%")
```

## 5. Which district has the lowest graduation rate?

```{r}
lowest_grad_rate_district <- districts %>% 
  arrange(grad) %>% 
  slice_head()

system_name <- lowest_grad_rate_district %>% select(system_name)
grad <- lowest_grad_rate_district %>% select(grad)

str_interp("The ${system_name} had the lowest graduation rate at ${grad}%.")
```

## 6. Which district has the highest dropout rate?

```{r}

highest_dropout <- districts %>% 
  arrange(desc(dropout)) %>% 
  slice_head()

system_name <- highest_dropout %>% 
  select(system_name)

dropout <- highest_dropout %>% 
  select(dropout)

str_interp("The ${system_name} had the highest dropout rate of ${dropout}%.")

```

## 7. Within the Mid Cumberland region, which district has the highest ACT composite?

```{r}

highest_act <- districts %>% 
  filter(region == "Mid Cumberland") %>% 
  arrange(desc(act_composite)) %>% 
  slice_head()

system_name <- highest_act %>% 
  select(system_name)

act <- highest_act %>% 
  select(act_composite)

str_interp("The ${system_name} had the highest act composite of ${act}%.")

```


## 8. Create a histogram showing the distribution of graduation rates. What can you say about this distribution?

```{r}

districts %>% 
  ggplot(aes(x = grad)) + 
  geom_histogram()

```

#### The majority of districts have graduation rates above 90%, there are several outliers with graduation rates below 75%.

##  9. Create a scatter plot to compare alg_1 proficiency rates to alg_2 rates. What do you notice? Facet this plot by region. Does anything stand out when you facet the plots?

```{r}

districts %>% 
  ggplot(aes(x = alg_1, y = alg_2, color = region)) + 
  geom_point() +
  geom_smooth(method = lm) +
  facet_wrap(~ region)

```

#### Upper cumberland increases in alg_1 scores do not show a corresponding increase in alg_2 scores. Southwest/Memphis had the largest spread in scores.


## 10. Create a bar chart showing the total enrollment by region. Which region has the highest total enrollment? Which has the smallest?

```{r}

districts %>% 
  group_by(region) %>% 
  ggplot(aes(x = reorder(region, enrollment, FUN=sum), y = enrollment, fill = region)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
  
```

## 11. When creating this bar chart you may have noticed that some districts have missing enrollment values. For how many districts is this the case?

```{r}

districts %>% 
  filter(is.na(enrollment)) %>% 
  count()

```

##  12. What is the mean graduation rate across all districts? What might be wrong with using just the regular mean to assess average graduation rates?

```{r}

districts %>% 
  summarise(mean_grad = mean(grad, na.rm = TRUE))
```


## 13. Redo the previous question but use a weighted average (`weighted.mean`) graduation across all districts, weighing by enrollment. How much does this change your answer? Can you explain using the data the reason for the big change from using the mean?

```{r}
districts %>% 
  summarise(mean_grad = weighted.mean(grad, enrollment, na.rm = TRUE))
```

## 14. Create a boxplot showing enrollment rates per region. Does anything stand out?

```{r}

districts %>% 
  group_by(region) %>% 
  ggplot(aes(color=region, x=region, y=enrollment)) +
  geom_boxplot(outlier.shape=NA) +
  scale_y_continuous(limits = c(0, 45000)) +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5))
```
There are some major metropolitan areas (Memphis, Nashville, etc) that are outliers with tens of thousands of students, while most TN districts have ~2500-5000 students. The boxplot above hides these outliers to improve readability.

## 15. Create a boxplot showing graduation rates per region. Does anything stand out?

```{r}
districts %>% 
  group_by(region) %>% 
  ggplot(aes(color=region, x=region, y=grad)) +
  geom_boxplot() + 
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5))
```
The median graduation rate seems to be between 85-95% for most districts, but there are some troubling outliers in some regions (Mid Cumberland, Southwest/Memphis, Northwest) where graduation rates are below 75%, sometimes well below.


## 16. Find the weighted average of graduation rates by region using enrollment as weights. Compare the results you get for the weighted average to what you see from the boxplots. Can you explain any discrepancy?

```{r}
districts %>% 
  group_by(region) %>%
  filter(!is.na(enrollment), !is.na(grad)) %>% 
  summarize(weighted_graduation_mean = weighted.mean(grad, enrollment)) %>% 
  arrange(desc(weighted_graduation_mean))
```

To me the results seem to be mostly align with the box plots.

## 17. For many districts, values for `alg_2` are lower than for `alg_1`. 

### 17a. Create a histogram showing the distribution of differences (`alg_1` - `alg_2`). 

```{r}
districts %>% 
  mutate(alg_diff=alg_1 - alg_2) %>% 
  group_by(region) %>% 
  ggplot(aes(color=region, x=alg_diff)) +
  geom_histogram()
```

### 17b. Which school had the largest drop from `alg_1` to `alg_2`? 

```{r}
districts %>% 
  mutate(alg_diff=alg_1 - alg_2) %>% 
  slice_max(n=5, order_by=alg_diff)
```

### 17c. For what percentage of schools is it true that `alg_2` is larger than `alg_1`? 

```{r}
districts %>%
  mutate(is_alg_2_larger=alg_2 > alg_1) %>% 
  count(is_alg_2_larger) %>% 
  filter(is_alg_2_larger == TRUE) %>% 
  pull(n) / count(districts %>% filter(!is.na(alg_1), !is.na(alg_2))) * 100
```

### 17d. Is there a similar dropoff for `eng_2` and `eng_3`?

```{r}
districts %>%
  mutate(is_eng_3_larger=eng_3 > eng_2) %>% 
  count(is_eng_3_larger) %>% 
  filter(is_eng_3_larger == TRUE) %>% 
  pull(n) / count(districts %>% filter(!is.na(eng_2), !is.na(eng_3))) * 100

```
Less than 1% of schools had higher eng_3 vs eng_2 rates, whereas over 12% of schools had higher alg_2 vs alg_1 rates.


## 18. You may have noticed that a lot of rows are missing values. Which district has the largest number of missing values? What do you notice about schools that have a lot of missing values?

```{r}
districts %>% 
  mutate(na_count = rowSums(is.na(.))) %>% 
  slice_max(n=5, order_by=na_count) %>% 
  group_by(region) %>% 
  ggplot(aes(fill=region, color=region, x=region, y=enrollment)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 90, hjust=1, vjust=0.5))
```
Mid Cumberland region schools have the highest number of missing values.

## 19. Find the correlation between graduation rate and all other variables. Create a horizontal bar chart showing these correlations. Make sure that your plot is ordered by correlation values. What do you notice from these correlations?





















